---
title: "HANPP_graphs"
author: "Kaeli Mueller"
date: "10/8/2020"
output:
  html_document: default
  pdf_document: default
---

######## Make initial bar charts #########
This is for myself to practice making bar charts with ggplot

Download important packages
Who knows if I need all of these, I'll load them all just in case
```{r}
#install.packages("statsr")
#install.packages("GGally")
#install.packages("bestglm")
#install.packages("lmtest")
#install.packages("normtest")
#install.packages("emmeans")
#install.packages("hrbrthemes")


library(tidyverse)
library(statsr)
library(MuMIn)
library(dplyr)
library(GGally)
library(car)
library(bestglm)
library(stats)
library(boot)
library(lmtest)
library(MASS)
library(normtest)
library(emmeans)
library(dplyr)
library(ggplot2)
library(lmtest)
library(reshape)
library(viridis)
library(hrbrthemes)
library(tidyselect)
```

Import data
```{r}
cotton_pima <- read.csv("C://Users//Kaeli//Documents//HANPP//data//HANPP_general//Cottonpima_FINALHANPP.csv", header = T)

hanpp_land <- read.csv("C://Users//Kaeli//Documents//HANPP//data//HANPP_general//HANPP_DataSet_LandEstimates.csv", 
                       header = TRUE)

#make dataset smaller and manageable
hanpp_land_small <- hanpp_land[1:6,]
head(hanpp_land_small)

#First try
#x axis is different counties, y axis is the HANPP harvest in g/C, fill takes it from the corn.grain column
p <- ggplot(data=hanpp_land_small, aes(x = County, 
                                       y = Total.HANPP.Harvest, 
                                       fill = Corn.Grain)) + 
  geom_bar(stat='identity', color = "black")

p
```
Now practice with a stacked bar chart
```{r}
#make dataset smaller, keep only the columns you want
hanpp_crop <- data.frame(hanpp_land_small$County, hanpp_land_small$Corn.Grain.1, hanpp_land_small$Cotton.Upland.1, hanpp_land_small$Soybeans.1, hanpp_land_small$Hardwood.1)

hanpp_crop_reshape <- melt(hanpp_crop, id =c("hanpp_land_small.County"))

head(hanpp_crop_reshape)

#rename
names(hanpp_crop_reshape)[1:3] <- c("County", "Crop", "HANPP_Harvest")

ggplot(data = hanpp_crop_reshape, aes(fill = Crop, 
                                      y = HANPP_Harvest, 
                                      x = County)) + 
  geom_bar(position = "stack", stat ="identity")
```

Fix some things to make the chart better
Change legend names (by changing the factor level names in the dataframe)
```{r}
#make them into character instead of factor
hanpp_crop_reshape$Crop <- as.character(hanpp_crop_reshape$Crop)

#rename all the weird labels into readable names
hanpp_crop_reshape[hanpp_crop_reshape == "hanpp_land_small.Corn.Grain.1"] <- "Corn"
hanpp_crop_reshape[hanpp_crop_reshape == "hanpp_land_small.Cotton.Upland.1"] <- "Upland Cotton"
hanpp_crop_reshape[hanpp_crop_reshape == "hanpp_land_small.Soybeans.1"] <- "Soybeans"
hanpp_crop_reshape[hanpp_crop_reshape == "hanpp_land_small.Hardwood.1"] <- "Hardwood"

#Get rid of random new column
hanpp_crop_final <- data.frame(hanpp_crop_reshape$County, hanpp_crop_reshape$Crop, hanpp_crop_reshape$HANPP_Harvest)

#Rename variable titles
names(hanpp_crop_final)[1:3] <- c("County", "Crop", "HANPP_Harvest")

#Coerce back to factor to 're-source' new factors
hanpp_crop_final$Crop <- factor(hanpp_crop_final$Crop)

print(hanpp_crop_final)

#Plot this new stacked bar chart
ggplot(data = hanpp_crop_final, aes(fill=Crop, 
                                    y = HANPP_Harvest, 
                                    x = County)) + 
  geom_bar(position = "stack", stat = "identity")
```

Customize the bar chart colors
```{r}
ggplot(data = hanpp_crop_final, aes(fill = Crop, 
                                    y = HANPP_Harvest, 
                                    x = County)) + 
  geom_bar(position = "stack", stat = "identity") + 
  scale_fill_manual(values = c("#FFD700","#663300", "#006400", "#ff0000"))
```

Improve the titles and axis lables
```{r}
ggplot(data = hanpp_crop_final, aes(fill=Crop, 
                                    y = HANPP_Harvest, 
                                    x = County)) + 
  geom_bar(position = "stack", stat = "identity") + 
  scale_fill_manual(values = c("#FFD700","#663300", "#006400", "#ff0000")) + 
  ggtitle("HANPP Harvest in Alabama Counties") + 
  ylab("HANPP Harvest (g C/m^2/year)")
```
That looks good, we have the basics of how we'll want our ggplot to look


######## Make bar graphs with the master excel sheet ########
Now we use the real data from whole excel sheet and format it to look like this 
```{r}
#try with a single county
#read in the entire csv
hanpp_main <- read.csv("C://Users//Kaeli//Documents//HANPP//data//HANPP_general//Master_HANPP_no_units.csv", header = TRUE)
head(hanpp_main)
#it has way too many columns we need to only filter what we need

#we eventually want to get it to work with every county
#But for now we'll just use one county

#filter out the county that you want
ALABAMA_AUTAUGA_2012 <- hanpp_main %>% 
  filter(State_County == "ALABAMA_AUTAUGA")

head(ALABAMA_AUTAUGA_2012)

#This selection was too much, narrow down the actual columns you need
#Take out columns only in the units of g m^2

head(ALABAMA_AUTAUGA_2012)

#X.8 = total used, X.2 = total hanpp, X.9 = unused, X.10 = above ground, X.11 = belowground

ALABAMA_AUTAUGA_2012 <- ALABAMA_AUTAUGA_2012 %>% 
  dplyr::select(Corn.Grain.2, Corn.Silage.2, Winter.Wheat.2, Spring.Wheat.Durum.2, Spring.Wheat.Excluding.Durum.2, Soybeans.2, Hay.Alfalfa.2, Cotton.Pima.2, Cotton.Upland.2, Sorghum.2, Other.Crops.2, Total.Crops.2, Hardwood.2, Softwood.2, Total.Timber.2, BLM.Grazing.2, USFS.Grazing.2, Private.Grazing.Land.2, Total.Grazing.2, X.2, X.8, X.9, X.10, X.11)

head(ALABAMA_AUTAUGA_2012)
#good now we have one county and all the columns that we need
```

Now we need to fix this data
```{r}
#rename the columns
#Give them readable names because these will show up on the final plot
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Corn.Grain.2"] <- "Corn Grain"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Corn.Silage.2"] <- "Corn Silage"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Winter.Wheat.2"] <- "Winter Wheat"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Spring.Wheat.Durum.2"] <- "Spring Wheat Dururm"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Spring.Wheat.Excluding.Durum.2"] <- "Spring Wheat Excluding Durum"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Soybeans.2"] <- "Soybeans"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Hay.Alfalfa.2"] <- "Hay Alfalfa"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Cotton.Pima.2"] <- "Cotton Pima"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Cotton.Upland.2"] <- "Cotton Upland"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Sorghum.2"] <- "Sorghum"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Other.Crops.2"] <- "Other Crops"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Total.Crops.2"] <- "Total Crops"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Hardwood.2"] <- "Hardwood"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Softwood.2"] <- "Softwood"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Total.Timber.2"] <- "Total Timber"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "BLM.Grazing.2"] <- "BLM Grazing"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "USFS.Grazing.2"] <- "USFS Grazing"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Private.Grazing.Land.2"] <- "Private Grazing Land"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "Total.Grazing.2"] <- "Total Grazing"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "X.2"] <- "Total HANPP Harvest"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "X.8"] <- "HANPP Harvest Used"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "X.9"] <- "HANPP Harvest Unused"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "X.10"] <- "HANPP Harvest Aboveground"
names(ALABAMA_AUTAUGA_2012)[names(ALABAMA_AUTAUGA_2012) == "X.11"] <- "HANPP Harvest Belowground"

#check to see if this worked
head(ALABAMA_AUTAUGA_2012)

#Make a column for types of HANPP and add it (this will be the x axis)

#Add the different values into the column

#We need a column for the x axis (different names of what we want each of the columns to be)
#a column for the different sections in each stacked bar chart, aka all of the current columns have to go into one column
#and a column for each of the numerical measurements of HANPP in gC.m2.year

```

#Try again solution
```{r}
hanpp_main <- read.csv("C://Users//Kaeli//Documents//HANPP//data//HANPP_general//Master_HANPP_no_units.csv", header = TRUE)

#select columns of interest
#
```


#### Cache valley
I could not figure out how to do this with the data so instead I got one county and manually made a csv
that looks like how I eventually want each county to look
```{r}
#read in the data
hanpp_cache <- read.csv("C://Users//Kaeli//Documents//HANPP//data//HANPP_general//bar_chart_test_cache.csv", header = TRUE)

#check if it's correct
head(hanpp_cache)
```
#Make the plots
```{r}
ggplot(data = hanpp_cache, aes(fill=Sections, 
                               y = gC.m2.year, 
                               x = Types_of_HANPP)) + 
  geom_bar(position = "stack", stat = "identity", show.legend = FALSE) + 
  #scale_fill_manual(values = c("#FFC300","#D06A0F", "#F1A25B", "#4AB617", "#715005", "#C8C7C4", "#6F6F6E", "#424140", "", )) +
  ggtitle("HANPP Harvest in Cache County") + 
  ylab("HANPP Harvest (g C/m^2/year)")

#why are the bars out of order?
#How to know how to assign the correct colors?
#But it's looking closer to what I need
```

